Dark Control:\\ A Unified Account of Default Mode Function\\ by Control Theory and Reinforcement Learning

### 2017-04-27 paper/paper.tex
stuff

The PMC matures relatively late (i.e., myelination) during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication \citep{flechsig1920}.
Normal and disturbed metabolic fluctuations in the
human PMC have been closely related to
changes of conscious awareness \citep{cavanna2006precuneus}.
%
Indeed,
the PMC has long been speculated to reflect constant computation of
environmental statistics and its internal representation
as an inner ``mind's eye'' \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
inability to combine various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as an inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity analyses
\citep{bzdok2015subspecialization}
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological significance assessment.

@@

\subsection{Sentinel hypothesis}
DMN regions have been associated with processing the
experienced or expected relevance of environment stimuli
\citep{montague2006imaging}.
Processing self-relevant information was perhaps the first
cognitive account that was proposed for DMN function \citep{gusnard2001medial}.
Since then,
many investigators have speculated that neural activity in the DMN
may reflect the brain's continuous tracking of
relevance in the environment
as an advantageous evolutionary adaptation \citep{randy2008, hahn2007cingulate}.
According to this cognitive account, the human brain's baseline realizes
a ``radar'' function to
detect subjectively relevant cues and unexpected events in the environment.
Propositions of a sentinel function to underlie DMN activity
have however seldom detailed
the mechanisms of
how attention and memory resources are exactly reallocated when
encountering a self-relevant environmental stimulus.
In the present formal account,
promising action trajectories
are recursively explored by human agents. Conversely,
certain branches of candidate action trajectories
are detected to be less worthy to become mentally explored.
This mechanism, expressed by the Bellman equation,
directly implies stratified allocation of attention and working memory load
over relevant cues and events in the environment.
%
Further,
our account provides a parsimonious explanation for
the consistently observed DMN implication
in certain goal-directed experimental tasks and
in task-unconstrained mind-wandering \citep{smith2009, bzdok2016formal}.
Both environment-detached and environment-engaged
cognitive processes may entail DMN recruitment
if real or imagined experience is processed, manipulated, and
used for predicting the future.
During tasks,
the policy and value estimates may be
updated to optimize especially short-term action.
At rest, these parameter updates may
improve especially mid- and long-term action.
The horizon is expressed in the $\gamma$ parameter in the formal account.
We thus provide answers for the currently unsettled question why the involvement
of the same neurobiological brain circuit (i.e., DMN) has been documented
for baseline house-keeping functions and specific task performances.

>>>>

The PMC matures relatively late (i.e., myelination) during postnatal development in monkeys
\citep{goldman1987development}, which is generally considered to
be a sign of evolutionary sophistication \citep{flechsig1920}.
Normal and disturbed metabolic fluctuations in the
human PMC have been closely related to
changes of conscious awareness \citep{cavanna2006precuneus}.
%
Indeed,
the PMC has long been speculated to reflect constant computation of
environmental statistics and its internal representation
as an inner ``mind's eye'' \citep{cavanna2006precuneus, leech_pcc2014}.
For instance, B\'alint's syndrome is a neurological disorder of conscious
awareness that results from medial damage in the parietal cortex
\citep{balint1909seelenlahmung}.
Patients are plagued by an
inability to combine various individual features of the visual
environment into an integrated whole (i.e., simultanagnosia)
as well as an inability to direct action towards
currently unattended environmental objects
(i.e., optic ataxia).
This can be viewed as a high-level impairment in gathering
information about alternative objects (i.e., exploration) as well as
leveraging these environmental opportunities (i.e., exploitation).
Congruently,
the human PMC was coupled in two functional connectivity analyses
\citep{bzdok2015subspecialization}
with the amygdala, involved in significance evaluation, and
the nucleus accumbens (NAc), involved in reward evaluation.
Specifically, among all parts of the PMC,
the ventral posterior cingulate cortex was
most connected to the laterobasal
nuclei group of the amygdala
\citep{bzdok2015subspecialization}.
This amygdalar subregion has been proposed to
continuously scan environmental input
for biological significance assessment.

@@

\subsection{Sentinel hypothesis}
DMN regions have been associated with processing the
experienced or expected relevance of environment stimuli
\citep{montague2006imaging}.
Processing self-relevant information was perhaps the first
cognitive account that was proposed for DMN function
\citep{gusnard2001medial, raichle2001pnas}.
Since then,
many investigators have speculated that neural activity in the DMN
may reflect the brain's continuous tracking of
relevance in the environment, such as spotting predators,
as an advantageous evolutionary adaptation \citep{randy2008, hahn2007cingulate}.
According to this cognitive account, the human brain's baseline realizes
a ``radar'' function to
detect subjectively relevant cues and unexpected events in the environment.
Propositions of a sentinel function to underlie DMN activity
have however seldom detailed
the mechanisms of
how attention and memory resources are exactly reallocated when
encountering a self-relevant environmental stimulus.
In the present formal account,
promising action trajectories
are recursively explored by human agents. Conversely,
certain branches of candidate action trajectories
are detected to be less worthy to become mentally explored.
This mechanism, expressed by the Bellman equation,
directly implies stratified allocation of attention and working memory load
over relevant cues and events in the environment.
%
Further,
our account provides a parsimonious explanation for
the consistently observed DMN implication
in certain goal-directed experimental tasks and
in task-unconstrained mind-wandering \citep{smith2009, bzdok2016formal}.
Both environment-detached and environment-engaged
cognitive processes may entail DMN recruitment
if real or imagined experience is processed, manipulated, and
used for predicting the future.
During tasks,
the policy and value estimates may be
updated to optimize especially short-term action.
At rest, these parameter updates may
improve especially mid- and long-term action.
The horizon is expressed in the $\gamma$ parameter in the formal account.
We thus provide answers for the currently unsettled question why the involvement
of the same neurobiological brain circuit (i.e., DMN) has been documented
for baseline house-keeping functions and specific task performances.
